---
title: "Hypothesis Test with The Bootstrap"
author: "Xinyi Lu"
date: "3/1/2022"
output: html_document
---

# Topic: Hypothesis Test with The Bootstrap
##Data set from surveys of customers of 168 Italian restaurants in New York City.
##The variables are:
## Price = the price (in $ US) of dinner (including 1 drink and tip)
## Food = customer rating of the food (out of 30)
## Decor = customer rating of the decor (out of 30)
## Service = customer rating of the service (out of 30)
## East = dummy variable, 1 (0) if the restaurant is east (west) of Fifth Avenue


## Project Question:
## Claim: Restaurants on East Fifth Avenue have higher food ratings than those on the west side.

$H_0: Food_{East} = Food_{West}$

$H_1: Food_{East} > Food_{West}$



```{r}
# loading package
library(tidyverse)
library(dplyr)
```

```{r}
# read cvs from Prof.Fox's github data set
nyc <- read.csv("https://ericwfox.github.io/data/nyc.csv")
str(nyc)
```
```{r}
#data wrangling
East <- nyc %>% filter(East == 1) %>% dplyr::select(Food)
West <- nyc %>% filter(East == 0) %>% dplyr::select(Food)
```

```{r}
str(East)
```

## Note: East and West are data frame

```{r}
length(East$Food)
length(West$Food)
```

```{r}
hist(West$Food)
```

```{r}
hist(East$Food)

```

## Applied Permutation Test and Hypothesis Test to restaurant dataset on East and West restaurants’ food rating means


```{r}
t.test(East$Food, West$Food, alternative = "greater", var.equal = TRUE)
```
## p-value = 0.009651 < alpha, when alpha = 0.05 



```{r}
# Permutation Test 
obv.mean <- mean(East$Food) - mean(West$Food)
obv.mean
```

```{r}
comb <- c(East$Food,West$Food)
diff.means <- numeric()

set.seed(234)
for (i in 1:5000) {
  means <- sample(comb, 168, replace = FALSE)
  diff.means[i] <- mean(means[1:106]) - mean(means[107:168])
}
```

```{r}
hist(diff.means)
```

```{r}
length(diff.means[ diff.means >= obv.mean]) / 5000
```


```{r}
# HT
# comb <- c(East$Food,West$Food)
# obv.mean <- mean(East$Food) - mean(West$Food) = 0.7388923
set.seed(345)
boot_sample <- matrix(sample(comb, 168 *10000, replace = TRUE), nrow = 10000)

#use apply() to return a vector
# boot_sample is a matrix
# margin 1: selecting by row
diff_means <- apply(boot_sample, 1, 
                    function(x){mean(x[1:106]) - mean(x[107:168])})
# ASL hat _ boot = number of mean difference 
length(diff_means[diff_means>= obv.mean])/10000

```

```{r}
# HT with studentized statistics
stud.t <- function(x){
  num <- mean(x[1:106]) - mean(x[107:168])
  pool.var <- (var(x[1:106])*105 + var(x[107:168])*61)/166
  denominator <- sqrt(pool.var*(1/106 + 1/62))
  return(num/denominator)
}

obv.stud.t <- stud.t(comb)

new.t <- apply(boot_sample, 1, stud.t)

length(new.t[new.t>= obv.stud.t])/10000
```



```{r}
boxplot(Food ~ East, data = nyc, ylab = "Food rating", xlab = "East (1 = East of Fifth Avenue")
```

